Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-10-01 psy_peds_3.Rmd
ratio estimation and visualization added

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"weeks" #possible values: days, weeks, months
```

@@

```{r}
thpd_psyPer <- count_icd %>%
  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer
```

>>>>

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

@@

##### Percentage 
```{r}
thpd_psyPer <- count_icd %>%
  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer
```
##### Ratio: patients with psychiatric condition / patients without psychiatric condition

```{r}
all_patients <- unique( all_codes_desc[, c("patient_num", "time_p")])
patients_with_psy_conditions <- unique( all_codes_psy[, c("patient_num", "time_p")] )
patients_without_psy_conditions <- setdiff(all_patients,patients_with_psy_conditions)

patients_without_psy_conditions_counts <- patients_without_psy_conditions %>%
  group_by( time_p ) %>%
  summarise(count_no_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic"))

patients_with_psy_conditions_counts <-  patients_with_psy_conditions %>%
  group_by( time_p ) %>%
  summarise(count_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic"))

ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
  mutate( ratio = count_psy_patients/count_no_psy_patients)

ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

ratio_psy_non_psy
